Alien species

### 2023-03-22 DRAFT_alien_species.Rmd
analysi of data area coverage

@@

I now have `r length(unique(HIA_muni$muni_HIF))` unique areas (for each main ecosystem) that I will, given
there is data, extrapolate indicator values over.

###### Subset ETs

I will subset the `alien_data` into the three ecosystems. Note that
only for open wetland do we have good ecosystem delineation maps to base
a spatial averaging on.

```{r}
#make valid
alien_data <- st_make_valid(alien_data)

@@

#### Looking at the HIA

See Fig. \@ref(fig.HIF-region) for a closer look at the distribution of HIA classes across Norway.

@@

@@

>>>>

See Fig. \@ref(fig:HIF-region) for a closer look at the distribution of HIA classes across Norway.

@@

I now have `r length(unique(HIA_muni$muni_HIF))` unique areas (for each main ecosystem) that I will, given
there is data, extrapolate indicator values over. But first I want to validate the assumption that the HIF 
explains a considerable portion of the variation i the indicator values.

I will subset the `alien_data` into the three ecosystems.

```{r}
#make geometries valid 
alien_data <- st_make_valid(alien_data)

@@

@@

### Exploring the indicator values

Most HIAs have good condition:

```{r, fig.cap="Number of HIAs with >20 data points (sum of all three ecosystems)"}
alien_extr %>%
  filter(!is.na(w_mean)) %>%
  ggplot()+
  geom_histogram(aes(x = w_mean))+
  theme_bw()+
  labs(x = "Area weighted indicator value",
       y = "Number of HIAs with >20 data points")
```

#### Data coverage

Let's see how big proportion of Norway we ended up having data for. Ideally I would do this after masking the indicator maps with ecosystem delineations, 
but I can also get a good feel for the data coverage like this as well.

The area of Norway is 
```{r}
(total_area <- st_area(outline))
```

And here is the total and relative area of the indicator maps
```{r relative-area, fig.cap="Relative area cover of the indicator maps for the alien species indicator (i.e. value 1 means that all HIAs had > 20 data points, and 0.2 means 20% of them did). If assuming the ecosystems are equally common where they are mapped and where hey are not, this measure of data coverage is representative of the real indicator coverage after maksing the maps with ecosystem delineations."}
alien_extr %>%
  filter(!is.na(w_mean)) %>%
  mutate(m2 = st_area(.)) %>%
  group_by(eco) %>%
  summarise(sum_area = sum(m2)) %>%
  add_column(total_area = total_area) %>%
  mutate(relative_area = sum_area/total_area) %>%
  ggplot()+
  geom_bar(aes(x = eco, y = relative_area),
           stat = "identity")+
  theme_bw()
```

Figure \@ref(fig:relative-area) shows us that this indicator will provide indicator values for about 20% of the semi-natural ecosystem, and considerably less for the other two ecosystems. The data coverage is very sensitive to the threshold for minimum data points, set to >20 in this case. Also, with more nature type data accumulating over time, the data coverage will increase. Therefore this indicator is worth while, and is seen as considerably better than the alternative to discard the entire data set.

@@

4. Then, look at [Median Summer Temperature] for how to aggregate spatially tp accounting areas (regions)
